\huge{Comment on `The Limits of Cosmic Shear'}

### 2016-12-05 Limber_Comment.tex
few changes and comments

\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\green}[1]{{\color{green}{#1}}}
\newcommand{\blue}[1]{{\color{blue}{#1}}}
\newcommand{\x}{\vec{x}}
\renewcommand{\d}[0]{{\rm d}}

@@

\begin{abstract}
Kitching et al. (2016) examine the impact of the flat-sky and Limber approximations that can be used when constraining cosmological parameters from the measurement of weak gravitational lensing by large-scale structures, also known as cosmic shear.  We support the core message of this paper that the use of such approximations should be considered in the future analyses of high precision surveys such as Euclid, LSST and WFIRST.  In this comment, we argue that these approximations have negligible impact for current surveys and are not able to explain away the tension between the constraints from some current weak lensing surveys and Planck.  Reducing uncertainty and characterising bias in photometric redshift estimation is a critical task for the success of current and future surveys.  We highlight that sophisticated treatments of these uncertainties have already been explored in the literature and have been shown not to resolve the tension with Planck, in contrast to the message that Kitching et al. (2016) convey.

@@

The measurement of weak gravitational lensing by large scale structures provides a powerful cosmological probe of dark matter, dark energy and modifications to gravity.  As such it is the primary science goal of several current (KiDS, HSC, DES) and future (Euclid, LSST, WFIRST) large-scale surveys.   Interest in the results from these surveys is high as statistically significant deviations have been found between the cosmological parameter constraints from the CMB Planck experiment \citep{planck/cosmo:2015} in comparison to weak lensing constraints from both the Kilo-Degree Survey \citep[KiDS;][]{hildebrandt/etal:2016} and the Canada-France-Hawaii Telescope Lensing Survey \citep[CFHTLenS;][] {joudaki/etal:2016}.  If the source of this tension is not a result of so-far unconsidered sources of systematic errors in one or all experiments, extensions to the standard flat $\Lambda$CDM cosmological models need to be considered.  \citet{joudaki/etal:2017} have shown, for example, that the tension can be resolved with an evolving dark energy model.

@@

This comment is structured as follows.  In section~\ref{sec:theory} we clarify which brand of approximations, highlighted by \citet{kitching/etal:2016}, have been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017}.  We show that we are unable to replicate key results from \citet{kitching/etal:2016} and therefore do not expect these approximations to impact on the cosmological analysis of current surveys. In section~\ref{sec:cfhtlens} we compare the `one-parameter' cosmological re-analysis of \citet{kitching/etal:2016} with the full six-parameter analysis by \citet{kilbinger/etal:2013}, arguing that this particular non-tomographic data set is not in significant tension with Planck.  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published.

>>>>

\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\green}[1]{{\color{green}{#1}}}
\newcommand{\blue}[1]{{\color{blue}{#1}}}
\newcommand{\x}{\vec{x}}
\renewcommand{\d}[0]{{\rm d}}
\newcommand{\shahab}[1]{\textcolor{pink}{\bf #1}}

@@

\begin{abstract}
Kitching et al. (2016) examine the impact of the flat-sky and Limber approximations that can be used when constraining cosmological parameters from the measurement of weak gravitational lensing by large-scale structures, also known as cosmic shear.  We support the core message of this paper that the use of such approximations should be considered in the future analyses of high precision surveys such as Euclid, LSST and WFIRST.  \shahab{(not sure it is clear that they are important even for future surveys, when considered together vs not considered at all)} In this comment, we argue that these approximations have negligible impact for current surveys and are not able to explain away the tension between the constraints from some current weak lensing surveys and Planck.  Reducing uncertainty and characterising bias in photometric redshift estimation is a critical task for the success of current and future surveys.  We highlight that sophisticated treatments of these uncertainties have already been explored in the literature and have been shown not to resolve the tension with Planck, in contrast to the message that Kitching et al. (2016) convey \shahab{(do we perhaps want to remove the last clause "in contrast to the message..." to make the draft somewhat less Kitching-centric given the weakness of that analysis?)}.

@@

The measurement of weak gravitational lensing by large scale structures provides a powerful cosmological probe of dark matter, dark energy, and modifications to gravity.  As such it is the primary science goal of several current (KiDS, HSC, DES) and future (Euclid, LSST, WFIRST) large-scale surveys. Interest in the results from these surveys is high as statistically significant deviations have been found between the cosmological parameter constraints from the CMB Planck experiment \citep{planck/cosmo:2015} in comparison to weak lensing constraints from both the Kilo-Degree Survey \citep[KiDS;][]{hildebrandt/etal:2016} and the Canada-France-Hawaii Telescope Lensing Survey \citep[CFHTLenS;][] {joudaki/etal:2016}.  If the source of this tension is not a result of so-far unconsidered sources of systematic errors in one or all experiments, extensions to the standard flat $\Lambda$CDM cosmological models need to be considered.  \citet{joudaki/etal:2017} have shown, for example, that the tension can be resolved with an evolving dark energy model.

@@

This comment is structured as follows.  In section~\ref{sec:theory} we clarify which brand of approximations, highlighted by \citet{kitching/etal:2016}, have been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017}.  We show that we are unable to replicate key results from \citet{kitching/etal:2016} and therefore do not expect these approximations to impact on the cosmological analysis of current surveys. In section~\ref{sec:cfhtlens} we compare the `one-parameter' cosmological re-analysis of \citet{kitching/etal:2016} with the full six-parameter analysis by \citet{kilbinger/etal:2013}, arguing that this particular non-tomographic data set is not in significant tension with Planck (\shahab{not sure I agree with this clause, please see my email}).  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published.

### 2016-12-05 cfhtlens.tex
few changes and comments

The Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS) represented a major step forward for the field of weak gravitational lensing, in terms of improved accuracy in data reduction \citep{erben/etal:2013}, the implementation of gaussianised matched multi-band photometry \citep{hildebrandt/etal:2012}, cross-correlation clustering analysis between photometric redshift slices to verify tomographic redshift distributions \citep{benjamin/etal:2013}, accurate calibrated shape measurements \citep{miller/etal:2013} and a full suite of informative systematic tests to select a clean data set \citep{heymans/etal:2012}.    Since the public release of this survey in 2013, the community has continued to scrutinise and advance our understanding of CFHTLenS by identifying a number of areas where the analysis could improve:
\begin{itemize}
\item{\citet{choi/etal:2016} identified biases in the tomographic photometric redshift distributions using a more effective clustering analysis, in comparison to \citet{benjamin/etal:2013}, by incorporating newly overlapping spectroscopy from the Sloan Digital Sky Survey.  The CFHTLenS tomographic cosmological analysis was then revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis based on the results from \citet{choi/etal:2016}, discussed further in section~\ref{sec:photoz}.}
\item{\citet{asgari/etal:2016} used the stringent `COSEBI' statistic to identify significant non-lensing `B-mode' distortions when the CFHTLenS data was split into tomographic slices.}
\item{\citet{kuijken/etal:2015} showed that the CFHTLenS shear calibration corrections derived in \citet{miller/etal:2013} were underestimated as a result of an imperfect match between the galaxy populations in the data and image simulations.}
\item{\citet{fenechconti/etal:2016} demonstrated that the CFHTLenS data would have been subject to a weight bias that favours galaxies that are more intrinsically oriented with the point-spread function.}
\item{\citet{joudaki/etal:2016} updated the CFHTLenS covariance matrices using larger-box numerical simulations that were less subject to the lack power on large scales.}
\end{itemize}
All these advances in our understanding were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{hildebrandt/etal:2016} which reports at $2.3 \sigma$ tension with Planck.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised shape measurements and calibrations for the shear and photometric redshifts.

>>>>

The Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS) represented a major step forward for the field of weak gravitational lensing, in terms of improved accuracy in data reduction \citep{erben/etal:2013}, the implementation of Gaussianised matched multi-band photometry \citep{hildebrandt/etal:2012}, cross-correlation clustering analysis between photometric redshift slices to verify tomographic redshift distributions \citep{benjamin/etal:2013}, accurate calibrated shape measurements \citep{miller/etal:2013} and a full suite of informative systematic tests to select a clean data set \citep{heymans/etal:2012}.    Since the public release of this survey in 2013, the community has continued to scrutinise and advance our understanding of CFHTLenS by identifying a number of areas where the analysis could improve:
\begin{itemize}
\item{\citet{choi/etal:2016} identified biases in the tomographic photometric redshift distributions using a more effective clustering analysis, in comparison to \citet{benjamin/etal:2013}, by incorporating newly overlapping spectroscopy from the Sloan Digital Sky Survey.  The CFHTLenS tomographic cosmological analysis was then revisited by \citet{joudaki/etal:2016} in order to include a full redshift error analysis based on the results from \citet{choi/etal:2016}, discussed further in section~\ref{sec:photoz}.}
\item{\citet{asgari/etal:2016} used the stringent `COSEBI' statistic to identify significant non-lensing `B-mode' distortions when the CFHTLenS data was split into tomographic slices.}
\item{\citet{kuijken/etal:2015} showed that the CFHTLenS shear calibration corrections derived in \citet{miller/etal:2013} were underestimated as a result of an imperfect match between the galaxy populations in the data and image simulations.}
\item{\citet{fenechconti/etal:2016} demonstrated that the CFHTLenS data would have been subject to a weight bias that favours galaxies that are more intrinsically oriented with the point-spread function.}
\item{\citet{joudaki/etal:2016} updated the CFHTLenS covariance matrices using larger-box numerical simulations that were less subject to the lack power on large scales.}
\end{itemize}
All these advances in our understanding were incorporated and accounted for in the recent KiDS cosmic shear analysis \citep{hildebrandt/etal:2016} which reports at $2.3 \sigma$ tension with Planck.  Efforts are now underway to fully re-analyse CFHTLenS using the advanced KiDS analysis pipeline with revised shape measurements and calibrations for the shear and photometric redshifts.

### 2016-12-05 theory.tex
added a red comment to fig 1
few changes and comments

\subsection{The Flat-Sky approximation and Extended Limber Approximation}
The baseline measurement for many cosmic shear studies has focused on the two-point shear correlation function between two tomographic bins $ij$ \citep[for more details see][and references therein]{bartelmann/schneider:2001}, given by
\be
\xi_\pm^{ij}(\theta) = \frac{1}{2\pi}\int \d\ell \,\ell \,P^{ij}_\kappa(\ell) \, J_{0,4}(\ell \theta) \, , 
\label{eqn:xiGG}
\ee
where $J_{0,4} (\ell \theta)$ is the zeroth (for $\xi_+$) or fourth (for $\xi_- $) order Bessel function of the first kind and $P_\kappa(\ell)$ is the convergence power spectrum at angular wave number $\ell$ which can be related to the underlying matter power spectrum $P_\delta$, using a Limber approximation as,
\be 
P^{ij}_\kappa(\ell) = T_l \int_0^{\chi_{\rm H}} \d \chi \, \frac{q_i(\chi)q_j(\chi)}{[f_K(\chi)]^2} \, P_\delta \left( \frac{\nu}{f_K(\chi)},\chi \right).
\label{eqn:Pkappa} 
\ee
Here $f_K(\chi)$ is the comoving angular diameter distance out to comoving radial distance $\chi$ and $\chi_{\rm H}$ is the comoving horizon distance.  The lensing efficiency function $q(\chi)$ is given in equation (5) of \citet{hildebrandt/etal:2016}.   \citet{kitching/etal:2016} show that the pre-factor $T_\ell$ is given by
\be
T_\ell = \frac{(\ell+2)(\ell+1)\ell(\ell-1)}{(\ell + 0.5)^4} \, .
\label{eqn:Tl}
\ee
When a survey is sufficiently small, a flat-sky approximation can be made reducing $T_\ell \approx 1$.  The value of $\nu$ depends on whether a baseline Limber approximation, or the more accurate extended Limber approximation of \citet{loverde/afshordi:2008} is used.  In the analysis of \citet{kitching/etal:2016} it is not fully clear which combinations of  $T_\ell$ and $\nu$ are considered, so we review them all as summarised in Table~\ref{tab:Tl_nu}, also highlighting which combinations were used in recent cosmic shear analyses.

@@

@@

 \begin{figure}
 \begin{center}
 \includegraphics[width=0.5\textwidth]{figures/Cl_xi_comp.pdf}
 \label{fig:Cl_xi}
 \caption{\emph{Top:} Theoretical models of the $\kappa$ power spectrum (\emph{left}) and shear correlation function $\xi_+$ (\emph{right})for the different approximations discussed in \citet{kitching/etal:2016}. \emph{Bottom:} Deviations with respect to the full treatment (LSH).}
 \end{center}
 \end{figure}

>>>>

\subsection{The Flat-Sky approximation and Extended Limber Approximation}
The baseline measurement for many cosmic shear studies has focused on the two-point shear correlation function between two tomographic bins $ij$ \citep[for more details see][and references therein]{bartelmann/schneider:2001}, given by
\be
\xi_\pm^{ij}(\theta) = \frac{1}{2\pi}\int \d\ell \,\ell \,P^{ij}_\kappa(\ell) \, J_{0,4}(\ell \theta) \, , 
\label{eqn:xiGG}
\ee
where $J_{0,4} (\ell \theta)$ is the zeroth (for $\xi_+$) or fourth (for $\xi_- $) order Bessel function of the first kind and $P_\kappa(\ell)$ is the convergence power spectrum (\shahab{should we perhaps rename this as the shear power spectrum as only equal to convergence power spectrum in the flat-sky limit, which we will not always use?}) at angular wave number $\ell$ which can be related to the underlying matter power spectrum $P_\delta$, using a Limber approximation as,
\be 
P^{ij}_\kappa(\ell) = T_l \int_0^{\chi_{\rm H}} \d \chi \, \frac{q_i(\chi)q_j(\chi)}{[f_K(\chi)]^2} \, P_\delta \left( \frac{\nu}{f_K(\chi)},\chi \right).
\label{eqn:Pkappa} 
\ee
Here $f_K(\chi)$ is the comoving angular diameter distance out to comoving radial distance $\chi$ and $\chi_{\rm H}$ is the comoving horizon distance.  The lensing efficiency function $q(\chi)$ is given in equation (5) of \citet{hildebrandt/etal:2016}.   \citet{kitching/etal:2016} show that the pre-factor $T_\ell$ \shahab{for the shear power spectrum} is given by
\be
T_\ell = \frac{(\ell+2)(\ell+1)\ell(\ell-1)}{(\ell + 0.5)^4} \, .
\label{eqn:Tl}
\ee
\shahab{The corresponding pre-factor for the convergence power spectrum (along with the pre-factors for a range of other statistics such as the CMB lensing power spectrum and galaxy-galaxy lensing power spectrum) is given in \citet{jk12}.}
When a survey is sufficiently small, a flat-sky approximation can be made reducing $T_\ell \approx 1$.  The value of $\nu$ depends on whether a baseline Limber approximation, or the more accurate extended Limber approximation of \citet{loverde/afshordi:2008} is used.  In the analysis of \citet{kitching/etal:2016} it is not fully clear which combinations of  $T_\ell$ and $\nu$ are considered, so we review them all as summarised in Table~\ref{tab:Tl_nu}, also highlighting which combinations were used in recent cosmic shear analyses.

@@

%perhaps include Joudaki & Kaplinghat (2012) reference here instead:
%\shahab{We further note that \citet{kitching/etal:2016} were not the first to consider the pre-factor from the extended Limber approximation and all-sky formulation of the power spectrum. For example, the corresponding pre-factor for the convergence power spectrum, along with the pre-factors for a range of other statistics such as the CMB lensing power spectrum and galaxy-galaxy lensing power spectrum, were given in \citet{jk12}.}

@@

 \begin{figure}
 \begin{center}
 \includegraphics[width=0.5\textwidth]{figures/Cl_xi_comp.pdf}
 \label{fig:Cl_xi}
 \caption{\emph{Top:} Theoretical models of the $\kappa$ power spectrum (\emph{left}) and shear correlation function $\xi_+$ (\emph{right})for the different approximations discussed in \citet{kitching/etal:2016}. \emph{Bottom:} Deviations with respect to the full treatment (LSH).\red{ Fig 3 in Kitching et al. is showing $C_\ell$ not $\ell(\ell+1)/2\pi C_\ell$. Would be better to show the same thing here ($C_\ell$ instead of $\ell C_\ell$ with the same axes range).} }
 \end{center}
 \end{figure}
